---
title: "Model Exploration"
author: "Rebecca Howard"
date: "8/29/2022"
output: html_document
---

```{r setup, include = FALSE}
### Load libraries
library(readxl)
library(plyr)
library(tibble)
library(here)
library(ggplot2)
library(lubridate)
library(date)
library(dplyr)
library(maps)
library(mapdata)
library(mgcv)
library(fields)
library(colorspace)
source("functions/vis_gam_COLORS.R")
```

```{r, include = FALSE, echo = FALSE}
# Histogram of catches
CPUE_hist <- function(species_subset){
  par(mfrow = c(1,2))
  hist(log(species_subset$catch + 1),
       xlab = "log(catch + 1)",
       main = "Raw Data")
  sum(1 * (species_subset$catch == 0)) / nrow(species_subset)
  hist(log(species_subset$catch[species_subset$catch > 0] + 1),
    xlab = "log(catch + 1)",
    main = "Normalized Data")
}

# Gaussian GAMs
# Need to consider whether to include oxygen, restricts the data a lot
# Only really significant for shortbelly rockfish
catch_GAMs <- function(species_subset){
  base_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian),
                  data = species_subset[species_subset$lncpue > 0,])
  ts_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                  s(temperature, k = 4) + s(salinity, k = 4),
                data = species_subset[species_subset$lncpue > 0,])  
  #tso_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
  #                 s(temperature, k = 4) + s(salinity, k = 4) + s(oxygen, k = 4),
  #               data = species_subset[species_subset$lncpue > 0,])  
  ts_geog_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                       s(temperature, k = 4) + s(salinity, k = 4) + s(longitude, latitude, by = mean_temp),
                     data = species_subset[species_subset$lncpue > 0,])    
  #tso_geog_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
  #                      s(temperature, k = 4) + s(salinity, k = 4) + s(oxygen, k = 4) + 
  #                      s(longitude, latitude, by = mean_temp),
  #                    data = species_subset[species_subset$lncpue > 0,])    
  ts_phen_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                       s(temperature, k = 4) + s(salinity, k = 4) + s(julian, by = mean_temp),
                     data = species_subset[species_subset$lncpue > 0,])    
  #tso_phen_gam <- gam(lncpue ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
  #                      s(temperature, k = 4) + s(salinity, k = 4) + s(oxygen, k = 4) + 
  #                      s(julian, by = mean_temp),
  #                    data = species_subset[species_subset$lncpue > 0,])    
  gam_list <- list(base_gam, ts_gam, ts_geog_gam, ts_phen_gam)
  best_gam <- gam_list[[which.min(sapply(1:length(gam_list),
                                         function(x) min(gam_list[[x]]$aic)))]] 
  return_list <- list(gam_list, best_gam)
} 

# Tweedie GAMs
tweedie_GAMs <- function(species_subset){
  base_tw <- gam(catch + 1 ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian),
                 family = tw(link = "log"),
                 method = 'REML',
                 data = species_subset)
  ts_tw <- gam(catch + 1 ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                 s(temperature, k = 4) + s(salinity, k = 4),
               family = tw(link = "log"),
               method = 'REML',
               data = species_subset)   
  ts_geog_tw <- gam(catch + 1 ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                      s(temperature, k = 4) + s(salinity, k = 4) + s(longitude, latitude, by = mean_temp),
                    family = tw(link = "log"),
                    method = 'REML',
                    data = species_subset)    
  ts_phen_tw <- gam(catch + 1 ~ factor(year) + s(longitude, latitude) + s(bottom_depth, k = 4) + s(julian) +
                       s(temperature, k = 4) + s(salinity, k = 4) + s(julian, by = mean_temp),
                    family = tw(link = "log"),
                    method = 'REML',
                    data = species_subset)   
  gam_list <- list(base_tw, ts_tw, ts_geog_tw, ts_phen_tw)
  best_gam <- gam_list[[which.min(sapply(1:length(gam_list),
                                         function(x) min(gam_list[[x]]$aic)))]] 
  return_list <- list(gam_list, best_gam)
} 

# Calculate variable coefficient term
variable_coefficient <- function(gam, data, variable){
  preds <- predict(gam[[2]], type = 'terms', se.fit = T)
  pred_slope <- preds[[1]][, 7] / variable
  pred_slope_se <- 1.96 * preds[[2]][, 7]
  pred_slope_up <- (preds[[1]][, 7] + pred_slope_se) / variable
  pred_slope_low <- (preds[[1]][, 7] - pred_slope_se) / variable
  sign_slope_pos <- (1:length(pred_slope))[pred_slope_low > 0]
  sign_slope_neg <- (1:length(pred_slope))[pred_slope_up < 0]
  return(list(sign_slope_neg, sign_slope_pos, pred_slope))
}

variable_coefficient2 <- function(gam, data, variable){
  preds <- predict(gam[[1]][[5]], type = 'terms', se.fit = T) # selects the PDO model
  pred_slope <- preds[[1]][, 7] / variable
  pred_slope_se <- 1.96 * preds[[2]][, 7]
  pred_slope_up <- (preds[[1]][, 7] + pred_slope_se) / variable
  pred_slope_low <- (preds[[1]][, 7] - pred_slope_se) / variable
  sign_slope_pos <- (1:length(pred_slope))[pred_slope_low > 0]
  sign_slope_neg <- (1:length(pred_slope))[pred_slope_up < 0]
  return(list(sign_slope_neg, sign_slope_pos, pred_slope))
}

variable_coefficient_lengths <- function(gam, data, variable){
  preds <- predict(gam, type = 'terms', se.fit = T)
  pred_slope <- preds[[1]][, 7] / variable
  pred_slope_se <- 1.96 * preds[[2]][, 7]
  pred_slope_up <- (preds[[1]][, 7] + pred_slope_se) / variable
  pred_slope_low <- (preds[[1]][, 7] - pred_slope_se) / variable
  sign_slope_pos <- (1:length(pred_slope))[pred_slope_low > 0]
  sign_slope_neg <- (1:length(pred_slope))[pred_slope_up < 0]
  return(list(sign_slope_neg, sign_slope_pos, pred_slope))
}

# Plot smooth terms
plot_variable <- function(gam, covariate, bounds, variable, ylabel, yvalues){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  plot(gam[[2]],
       pages = 0,
       select = covariate, # 1 = year/PDO/NPGO, 2 = lat/lon, 3 = depth, 4 = julian, 5 = temp
       shade = T,
       shade.col = "lemonchiffon3",
       ylim = bounds,
       xlab = variable,
       ylab = ylabel,
       yaxt = yvalues,
       seWithMean = T,
       scale = 0,
       cex.axis = 3,
       cex.lab = 3,
       family = "serif",
       lwd = 2)
}

# Plot distribution
location_plot_lengths <- function(gam, species_subset) {
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam,
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7, -116.5),
            ylim = range(species_subset$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  maps::map('worldHires',
            add = T,
            col = 'antiquewhite4',
            fill = T)
  image.plot(legend.only = T,
           col = jet.colors(100),
           legend.shrink = 0.2,
           smallplot = c(.28, .31, .11, .24),
           legend.cex = 1.3,
           axis.args = list(cex.axis = 1.8,
                            family = "serif"),
           legend.width = 0.8,
           legend.mar = 6,
           zlim = c(min(gam$linear.predictors), 
                    max(gam$linear.predictors)),
           legend.args = list("log(cpue+1)",
                              side = 2, cex = 2,
                              family = "serif"))
}

location_plot <- function(gam, species_subset) {
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam[[2]],
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7, -116.5),
            ylim = range(species_subset$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  maps::map('worldHires',
            add = T,
            col = 'antiquewhite4',
            fill = T)
  image.plot(legend.only = T,
           col = jet.colors(100),
           legend.shrink = 0.2,
           smallplot = c(.28, .31, .11, .24),
           legend.cex = 1.3,
           axis.args = list(cex.axis = 1.8,
                            family = "serif"),
           legend.width = 0.8,
           legend.mar = 6,
           zlim = c(min(gam[[2]]$linear.predictors), 
                    max(gam[[2]]$linear.predictors)),
           legend.args = list("log(cpue+1)",
                              side = 2, cex = 2,
                              family = "serif"))
}

# Plot variable coefficient term
plot_var_coef <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam[[2]],
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7,-116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  symbols(data$lon[predictions[[2]]],
          data$lat[predictions[[2]]],
          circle = predictions[[3]][predictions[[2]]],
          inches = 0.12,
          add = T,
          bg = alpha('darkred', 0.4),
          fg = alpha('black', 0.08))
  symbols(data$lon[predictions[[1]]],
          data$lat[predictions[[1]]],
          circle = (-1) * predictions[[3]][predictions[[1]]],
          inches = 0.12,
          add = T,
          bg = alpha('navy', 0.4),
          fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam[[2]]$linear.predictors),
                      max(gam[[2]]$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

plot_var_coef2 <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam[[2]],
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7,-116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  symbols(data$lon[predictions[[2]]],
          data$lat[predictions[[2]]],
          circle = predictions[[3]][predictions[[2]]],
          inches = 0.12,
          add = T,
          bg = alpha('darkred', 0.4),
          fg = alpha('black', 0.08))
  # symbols(data$lon[predictions[[1]]],
  #         data$lat[predictions[[1]]],
  #         circle = (-1) * predictions[[3]][predictions[[1]]],
  #         inches = 0.12,
  #         add = T,
  #         bg = alpha('navy', 0.4),
  #         fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam[[2]]$linear.predictors),
                      max(gam[[2]]$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

# Plot variable coefficient term
plot_var_coef_PDO <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam[[1]][[5]],
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7,-116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  symbols(data$lon[predictions[[2]]],
          data$lat[predictions[[2]]],
          circle = predictions[[3]][predictions[[2]]],
          inches = 0.12,
          add = T,
          bg = alpha('darkred', 0.4),
          fg = alpha('black', 0.08))
  symbols(data$lon[predictions[[1]]],
          data$lat[predictions[[1]]],
          circle = (-1) * predictions[[3]][predictions[[1]]],
          inches = 0.12,
          add = T,
          bg = alpha('navy', 0.4),
          fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam[[2]]$linear.predictors),
                      max(gam[[2]]$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

plot_var_coef_PDO2 <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam[[1]][[5]],
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'response',
            xlim = c(-125.7,-116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  # symbols(data$lon[predictions[[2]]],
  #         data$lat[predictions[[2]]],
  #         circle = predictions[[3]][predictions[[2]]],
  #         inches = 0.12,
  #         add = T,
  #         bg = alpha('darkred', 0.4),
  #         fg = alpha('black', 0.08))
  symbols(data$lon[predictions[[1]]],
          data$lat[predictions[[1]]],
          circle = (-1) * predictions[[3]][predictions[[1]]],
          inches = 0.12,
          add = T,
          bg = alpha('navy', 0.4),
          fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam[[2]]$linear.predictors),
                      max(gam[[2]]$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

plot_var_coef_lengths <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam,
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7, -116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  symbols(data$lon[predictions[[2]]],
          data$lat[predictions[[2]]],
          circle = predictions[[3]][predictions[[2]]],
          inches = 0.12,
          add = T,
          bg = alpha('darkred', 0.4),
          fg = alpha('black', 0.08))
  symbols(data$lon[predictions[[1]]],
          data$lat[predictions[[1]]],
          circle = (-1) * predictions[[3]][predictions[[1]]],
          inches = 0.12,
          add = T,
          bg = alpha('navy', 0.4),
          fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam$linear.predictors),
                      max(gam$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

plot_var_coef_lengths2 <- function(gam, data, predictions){
  par(mar = c(6.4, 7.2, .5, 0.6) + 0.1,
      oma = c(1, 1, 1, 1),
      mgp = c(5, 2, 0))
  myvis_gam(gam,
            view = c('lon', 'lat'),
            too.far = 0.07,
            plot.type = 'contour',
            contour.col = contour_col,
            color = "jet" ,
            type = 'link',
            xlim = c(-125.7, -116.5),
            ylim = range(data$lat, na.rm = TRUE) + c(-.4, .5),
            family = "serif",
            xlab = "Longitude",
            ylab = "Latitude",
            main = " ",
            cex.lab = 2.5,
            cex.axis =  2.5)
  # symbols(data$lon[predictions[[2]]],
  #         data$lat[predictions[[2]]],
  #         circle = predictions[[3]][predictions[[2]]],
  #         inches = 0.12,
  #         add = T,
  #         bg = alpha('darkred', 0.4),
  #         fg = alpha('black', 0.08))
  symbols(data$lon[predictions[[1]]],
          data$lat[predictions[[1]]],
          circle = (-1) * predictions[[3]][predictions[[1]]],
          inches = 0.12,
          add = T,
          bg = alpha('navy', 0.4),
          fg = alpha('black', 0.08))
  map("worldHires",
      fill = T,
      col = "wheat4",
      add = T)
  image.plot(legend.only = T,
             col = jet.colors(100),
             legend.shrink = 0.2,
             smallplot = c(.28, .31, .11, .24),
             legend.cex = 1.3,
             axis.args = list(cex.axis = 1.8,
                              family = "serif"),
             legend.width = 0.8,
             legend.mar = 6,
             zlim = c(min(gam$linear.predictors),
                      max(gam$linear.predictors)),
             legend.args = list("log(cpue+1)",
                                side = 2, cex = 2,
                                family = "serif"))
}

contour_col <- rgb(0, 0, 255, max = 255, alpha = 0, names = "white")
jet.colors <- colorRampPalette(c(sequential_hcl(15, palette = "Mint")))

merge_data <- function(data1, data2){
  merge(data1, data2,
        by.x = c("cruise", "haul_no"),
        by.y = c("cruise", "haul_no"),
        all.x = F, all.y = T)
}
```

```{r}
# Load data
yoy_hake <- readRDS(file = "../data/yoy_hake.Rdata")
yoy_dab <- readRDS(file = "../data/yoy_sanddab.Rdata")
yoy_anch <- readRDS(file = "../data/yoy_anchovy.Rdata")
yoy_widow <- readRDS(file = "../data/yoy_widow.Rdata")
yoy_sbelly <- readRDS(file = "../data/yoy_shortbelly.Rdata")
RREAS_list <- readRDS(file = "../data/RREAS_list.Rdata")
ctd_means <- readRDS(file = "../data/RREAS_ctd_means.Rds")
```

```{r}
CPUE_hist(RREAS_list[[1]])
CPUE_hist(RREAS_list[[2]])
CPUE_hist(RREAS_list[[3]])
CPUE_hist(RREAS_list[[4]])
CPUE_hist(RREAS_list[[5]])
```

Match with CTD means
```{r}
yoy_hake$mean_temp <- ctd_means$temperature[match(yoy_hake$year, ctd_means$year)]
yoy_dab$mean_temp <- ctd_means$temperature[match(yoy_dab$year, ctd_means$year)]
yoy_anch$mean_temp <- ctd_means$temperature[match(yoy_anch$year, ctd_means$year)]
yoy_widow$mean_temp <- ctd_means$temperature[match(yoy_widow$year, ctd_means$year)]
yoy_sbelly$mean_temp <- ctd_means$temperature[match(yoy_sbelly$year, ctd_means$year)]

RREAS_list <- lapply(RREAS_list, function(x)
  mutate(x, mean_temp = ctd_means$temperature[match(x$year, ctd_means$year)],
         lncpue = log(x$catch + 1)))
```

### Gaussian on positive catches
## YOY Hake
```{r}
# Base GAM
hake_base <- gam(lncpue ~ factor(year) +
                   s(julian) +
                   s(longitude, latitude) +
                   s(bottom_depth, k = 4),
                 data = RREAS_list[[1]][RREAS_list[[1]]$lncpue > 0,])
summary(hake_base)

par(mfrow = c(2, 2))
gam.check(hake_base)

par(mfrow = c(2, 2))
plot(hake_base)


# Select best GAM
hake_gam <- catch_GAMs(RREAS_list[[1]])
summary(hake_gam[[2]])

par(mfrow = c(2, 2))
gam.check(hake_gam[[2]])

par(mfrow = c(2, 2))
plot(hake_gam[[2]])
```

## YOY Sanddab
```{r}
# Base GAM
dab_base <- gam(lncpue ~ factor(year) +
                  s(julian) +
                  s(longitude, latitude) +
                  s(bottom_depth, k = 4),
                data = RREAS_list[[2]][RREAS_list[[2]]$lncpue > 0,])
summary(dab_base)

par(mfrow = c(2, 2))
gam.check(dab_base)

par(mfrow = c(2, 2))
plot(dab_base)


# Select best GAM
dab_gam <- catch_GAMs(RREAS_list[[2]])
summary(dab_gam[[2]])

par(mfrow = c(2, 2))
gam.check(dab_gam[[2]])

par(mfrow = c(2, 2))
plot(dab_gam[[2]])
```


## YOY Anchovy
```{r}
# Base GAM
anch_base <- gam(lncpue ~ factor(year) +
                  s(julian) +
                  s(longitude, latitude) +
                  s(bottom_depth, k = 4),
                data = RREAS_list[[3]][RREAS_list[[3]]$lncpue > 0,])
summary(anch_base)

par(mfrow = c(2, 2))
gam.check(anch_base)

par(mfrow = c(2, 2))
plot(anch_base)


# Select best GAM
anch_gam <- catch_GAMs(RREAS_list[[3]])
summary(anch_gam[[2]])

par(mfrow = c(2, 2))
gam.check(anch_gam[[2]])

par(mfrow = c(2, 2))
plot(anch_gam[[2]])
```

## YOY Widow Rockfish
```{r}
# Base GAM
widow_base <- gam(lncpue ~ factor(year) +
                  s(julian) +
                  s(longitude, latitude) +
                  s(bottom_depth, k = 4),
                data = RREAS_list[[4]][RREAS_list[[4]]$lncpue > 0,])
summary(widow_base)

par(mfrow = c(2, 2))
gam.check(widow_base)

par(mfrow = c(2, 2))
plot(widow_base)


# Select best GAM
widow_gam <- catch_GAMs(RREAS_list[[4]])
summary(widow_gam[[2]])

par(mfrow = c(2, 2))
gam.check(widow_gam[[2]])

par(mfrow = c(2, 2))
plot(widow_gam[[2]])
```

## YOY Shortbelly Rockfish
```{r}
# Base GAM
sbelly_base <- gam(lncpue ~ factor(year) +
                  s(julian) +
                  s(longitude, latitude) +
                  s(bottom_depth, k = 4),
                data = RREAS_list[[5]][RREAS_list[[5]]$lncpue > 0,])
summary(sbelly_base)

par(mfrow = c(2, 2))
gam.check(sbelly_base)

par(mfrow = c(2, 2))
plot(sbelly_base)


# Select best GAM
sbelly_gam <- catch_GAMs(RREAS_list[[5]])
summary(sbelly_gam[[2]])

par(mfrow = c(2, 2))
gam.check(sbelly_gam[[2]])

par(mfrow = c(2, 2))
plot(sbelly_gam[[2]])
```

### Tweedie on counts
## YOY Hake
```{r}
# Base GAM
hake_base_t <- gam(catch + 1 ~ factor(year) +
                   s(julian) +
                   s(longitude, latitude) +
                   s(bottom_depth, k = 4),
                 family = tw(link = "log"),
                 method = 'REML',
                 data = RREAS_list[[1]])
summary(hake_base_t)

par(mfrow = c(2, 2))
gam.check(hake_base_t)

par(mfrow = c(2, 2))
plot(hake_base_t)


# Select best GAM
hake_gam_t <- tweedie_GAMs(RREAS_list[[1]])
summary(hake_gam_t[[2]])

AIC(hake_gam_t[[2]])

par(mfrow = c(2, 2))
gam.check(hake_gam_t[[2]])

par(mfrow = c(2, 2))
plot(hake_gam_t[[2]])
```

```{r}
pred_hake_t <- variable_coefficient(hake_gam_t, RREAS_list[[1]], RREAS_list[[1]]$mean_temp)

plot_var_coef(hake_gam_t, RREAS_list[[1]], pred_hake_t)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_var_coef.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```


```{r}
plot_variable(hake_gam_t, 
              2, 
              c(-3.5, 4.5), 
              "Depth", 
              "Species Abundance Anomalies", 
              "s")
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_depth.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
plot_variable(hake_gam_t, 
              3, 
              c(-3.5, 4.5), 
              "Day of Year", 
              "", 
              "n")
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_julian.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot(hake_gam_t,
              yoy_hake)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_distribution.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Variable coefficient term for PDO model
```{r}
pred_yoy_hake_t <- variable_coefficient2(hake_gam_t, yoy_hake, yoy_hake$PDO_pos)

plot_var_coef_PDO(hake_gam_t, yoy_hake, pred_yoy_hake_t)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_var_coef_PDO.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Size-specific models
```{r}
hake_lower <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_hake_sp, 
                        FUN = length, 
                        subset = std_length <= 30)
yoy_hake <- merge_data(hake_lower, yoy_hake)
names(yoy_hake)[names(yoy_hake) == "std_length"] <- "lower_num"

hake_upper <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_hake_sp, 
                        FUN = length, 
                        subset = as.numeric(std_length) > 30)
yoy_hake <- merge_data(hake_upper, yoy_hake)
names(yoy_hake)[names(yoy_hake) == "std_length"] <- "upper_num"

hake_total <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_hake_sp, 
                        FUN = length)
yoy_hake <- merge_data(hake_total, yoy_hake)
names(yoy_hake)[names(yoy_hake) == "std_length"] <- "total_num"

yoy_hake$upper_cpue <- (yoy_hake$upper_num / yoy_hake$total_num) * yoy_hake$catch
yoy_hake$lower_cpue <- (yoy_hake$lower_num / yoy_hake$total_num) * yoy_hake$catch

yoy_hake$upper_cpue[is.na(yoy_hake$upper_cpue)] <- 0
yoy_hake$lower_cpue[is.na(yoy_hake$lower_cpue)] <- 0

yoy_hake_filtered <- yoy_hake[yoy_hake$year > 1993, ] # no length data before 1994
```

```{r}
# Select best GAM
yoy_hake_gam_all <- tweedie_GAMs(yoy_hake_filtered)
summary(yoy_hake_gam_all[[2]])

par(mfrow = c(2, 2))
gam.check(yoy_hake_gam_all[[2]])

par(mfrow = c(2, 2))
plot(yoy_hake_gam_all[[2]])
```

```{r}
pred_yoy_hake_all <- variable_coefficient(yoy_hake_gam_all, yoy_hake_filtered, yoy_hake_filtered$NPGO_pos)

plot_var_coef(yoy_hake_gam_all, yoy_hake_filtered, pred_yoy_hake_all)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_var_coef_all.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot(yoy_hake_gam_all,
              yoy_hake_filtered)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_distribution_all.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Histogram of lengths
```{r}
hist(yoy_hake_sp$std_length)
```

Split into two models
```{r}
hake_lower <- gam(lower_cpue + 1 ~ factor(year) + s(lon, lat) + s(bottom_depth, k = 4) + s(julian) + s(longitude, latitude, by = NPGO_pos),
                  family = tw(link = "log"),
                  method = 'REML',
                  data = yoy_hake_filtered)
hake_upper <- gam(upper_cpue + 1 ~ factor(year) + s(lon, lat) + s(bottom_depth, k = 4) + s(julian) + s(longitude, latitude, by = NPGO_pos),
                  family = tw(link = "log"),
                  method = 'REML',
                  data = yoy_hake_filtered)
summary(hake_lower)
summary(hake_upper)
```

```{r}
location_plot_lengths(hake_lower,
              yoy_hake_filtered)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_lower_dis.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot_lengths(hake_upper,
              yoy_hake_filtered)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_upper_dis.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
pred_yoy_hake_lower <- variable_coefficient_lengths(hake_lower, yoy_hake_filtered, yoy_hake_filtered$NPGO_pos)

plot_var_coef_lengths(hake_lower, yoy_hake_filtered, pred_yoy_hake_lower)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_var_coef_lower.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
pred_yoy_hake_upper <- variable_coefficient_lengths(hake_upper, yoy_hake_filtered, yoy_hake_filtered$NPGO_pos)

plot_var_coef_lengths(hake_upper, yoy_hake_filtered, pred_yoy_hake_upper)
dev.copy(jpeg, here('results/yoy_hake_preliminary', 'yoy_hake_var_coef_upper.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```


## YOY Sanddab
```{r}
# Base GAM
yoy_dab_base_t <- gam(catch + 1 ~ factor(year) +
                      s(julian) +
                      s(longitude, latitude) +
                      s(bottom_depth, k = 4),
                    family = tw(link = "log"),
                    method = 'REML',
                    data = yoy_dab_catch)
summary(yoy_dab_base_t)

par(mfrow = c(2, 2))
gam.check(yoy_dab_base_t)

par(mfrow = c(2, 2))
plot(yoy_dab_base_t)


# Select best GAM
yoy_dab_gam_t <- tweedie_GAMs(yoy_dab_catch)
summary(yoy_dab_gam_t[[2]])

par(mfrow = c(2, 2))
gam.check(yoy_dab_gam_t[[2]])

par(mfrow = c(2, 2))
plot(yoy_dab_gam_t[[2]])
```

```{r}
pred_yoy_dab_t <- variable_coefficient(yoy_dab_gam_t, yoy_dab_catch, yoy_dab_catch$NPGO_pos)

plot_var_coef2(yoy_dab_gam_t, yoy_dab_catch, pred_yoy_dab_t)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_var_coef.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
plot_variable(yoy_dab_gam_t, 
              2, 
              c(-2.5, 1.5), 
              "Depth", 
              "Species Abundance Anomalies", 
              "s")
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_depth.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
plot_variable(yoy_dab_gam_t, 
              3, 
              c(-2.5, 1.5), 
              "Day of Year", 
              "", 
              "n")
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_julian.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot(yoy_dab_gam_t,
              yoy_dab_catch)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_distribution.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Variable coefficient term for PDO model
```{r}
pred_yoy_dab_t <- variable_coefficient2(yoy_dab_gam_t, yoy_dab_catch, yoy_dab_catch$PDO_pos)

plot_var_coef_PDO2(yoy_dab_gam_t, yoy_dab_catch, pred_yoy_dab_t)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_var_coef_PDO.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Length models
```{r}
# Filter based on years with length data
dab_lower <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_dab_sp, 
                        FUN = length, 
                        subset = std_length <= 37)
yoy_dab <- merge_data(dab_lower, yoy_dab_catch)
names(yoy_dab)[names(yoy_dab) == "std_length"] <- "lower_num"

dab_upper <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_dab_sp, 
                        FUN = length, 
                        subset = as.numeric(std_length) > 37)
yoy_dab <- merge_data(dab_upper, yoy_dab)
names(yoy_dab)[names(yoy_dab) == "std_length"] <- "upper_num"

dab_total <- aggregate(std_length ~ haul_no + cruise, 
                        data = yoy_dab_sp, 
                        FUN = length)
yoy_dab <- merge_data(dab_total, yoy_dab)
names(yoy_dab)[names(yoy_dab) == "std_length"] <- "total_num"

yoy_dab$upper_cpue <- (yoy_dab$upper_num / yoy_dab$total_num) * yoy_dab$catch
yoy_dab$lower_cpue <- (yoy_dab$lower_num / yoy_dab$total_num) * yoy_dab$catch

yoy_dab$upper_cpue[is.na(yoy_dab$upper_cpue)] <- 0
yoy_dab$lower_cpue[is.na(yoy_dab$lower_cpue)] <- 0

yoy_dab_filtered <- yoy_dab[yoy_dab$year > 2010, ] # Length data only from 2011 onward
```

```{r}
# Select best GAM
yoy_dab_gam_all <- tweedie_GAMs(yoy_dab_filtered)
summary(yoy_dab_gam_all[[2]])

par(mfrow = c(2, 2))
gam.check(yoy_dab_gam_all[[2]])

par(mfrow = c(2, 2))
plot(yoy_dab_gam_all[[2]])
```

```{r}
pred_yoy_dab_all <- variable_coefficient(yoy_dab_gam_all, yoy_dab_filtered, yoy_dab_filtered$NPGO_pos)

plot_var_coef2(yoy_dab_gam_all, yoy_dab_filtered, pred_yoy_dab_all)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_var_coef.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot(yoy_dab_gam_all,
              yoy_dab_filtered)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_distribution_all.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```
Histogram of lengths
```{r}
hist(yoy_dab_sp$std_length)
```

```{r}
dab_lower <- gam(lower_cpue + 1 ~ factor(year) + s(lon, lat) + s(bottom_depth, k = 4) + s(julian) + s(NPGO_pos),
                  family = tw(link = "log"),
                  method = 'REML',
                  data = yoy_dab_filtered)
dab_upper <- gam(upper_cpue + 1 ~ factor(year) + s(lon, lat) + s(bottom_depth, k = 4) + s(julian) + s(NPGO_pos),
                  family = tw(link = "log"),
                  method = 'REML',
                  data = yoy_dab_filtered)
summary(dab_lower)
summary(dab_upper)
```

```{r}
location_plot_lengths(dab_lower,
              yoy_dab_filtered)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_lower_dis.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot_lengths(dab_upper,
              yoy_dab_filtered)
dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_upper_dis.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

Use this chunk and one below if something changes to incorporate VC term
```{r}
# pred_yoy_dab_lower <- variable_coefficient_lengths(dab_lower, yoy_dab_filtered, yoy_dab_filtered$NPGO_pos)
# 
# plot_var_coef_lengths(dab_lower, yoy_dab_filtered, pred_yoy_dab_lower)
# dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_var_coef_lower.jpg'), 
#          height = 15, width = 8, units = 'in', res = 200)
# dev.off()
```

```{r}
# pred_yoy_dab_upper <- variable_coefficient_lengths(dab_upper, yoy_dab_filtered, yoy_dab_filtered$NPGO_pos)
# 
# plot_var_coef_lengths2(dab_upper, yoy_dab_filtered, pred_yoy_dab_upper)
# dev.copy(jpeg, here('results/yoy_dab_preliminary', 'yoy_dab_var_coef_upper.jpg'), 
#          height = 15, width = 8, units = 'in', res = 200)
# dev.off()
```

## Adult Sanddab
```{r}
# Base GAM
adult_dab_base_t <- gam(catch + 1 ~ factor(year) +
                        s(julian) +
                        s(longitude, latitude) +
                        s(bottom_depth, k = 4),
                      family = tw(link = "log"),
                      method = 'REML',
                      data = adult_dab_catch)
summary(adult_dab_base_t)

par(mfrow = c(2, 2))
gam.check(adult_dab_base_t)

par(mfrow = c(2, 2))
plot(adult_dab_base_t)


# Select best GAM
adult_dab_gam_t <- tweedie_GAMs(adult_dab_catch)
summary(adult_dab_gam_t[[2]])

par(mfrow = c(2, 2))
gam.check(adult_dab_gam_t[[2]])

par(mfrow = c(2, 2))
plot(adult_dab_gam_t[[2]])
```

```{r}
pred_adult_dab_t <- variable_coefficient(adult_dab_gam_t, adult_dab_catch, adult_dab_catch$PDO_pos)

plot_var_coef(adult_dab_gam_t, adult_dab_catch, pred_adult_dab_t)
dev.copy(jpeg, here('results/adult_dab_preliminary', 'adult_dab_var_coef.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```

```{r}
plot_variable(adult_dab_gam_t, 
              2, 
              c(-1.5, 0.5), 
              "Depth", 
              "Species Abundance Anomalies", 
              "s")
dev.copy(jpeg, here('results/adult_dab_preliminary', 'adult_dab_depth.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
plot_variable(adult_dab_gam_t, 
              3, 
              c(-1.5, 0.5), 
              "Day of Year", 
              "", 
              "n")
dev.copy(jpeg, here('results/adult_dab_preliminary', 'adult_dab_julian.jpg'), 
         height = 10, width = 10, units = 'in', res = 200)
dev.off()
```

```{r}
location_plot(adult_dab_gam_t,
              adult_dab_catch)
dev.copy(jpeg, here('results/adult_dab_preliminary', 'adult_dab_distribution.jpg'), 
         height = 15, width = 8, units = 'in', res = 200)
dev.off()
```
